---
title: "Basic Data Manipulation"
output: html_notebook
---

The data is in an Excel file here called refine.xlsx. Right away, you'll notice that the data set has a few issues:
* There are four brands: Philips, Akzo, Van Houten and Unilever. However, there are many different spellings and capitalizations of those names!
* The product code and number are combined in one column, separated by a hyphen.

```{r Library load, message=FALSE, warning=FALSE, include=FALSE}
install.packages("stringdist")
library("stringdist")
install.packages("dplyr")
library("dplyr")
install.packages("tidyr")
library("tidyr")
install.packages("dummies")
library("dummies")
```

```{r Load the data in RStudio, message=FALSE, warning=FALSE, include=FALSE}
dfOriginal <- read.csv("refine_original.csv")
df <- dfOriginal
```

Clean up the 'company' column so all of the misspellings of the brand names are standardized. For example, you can transform the values in the column to be: philips, akzo, van houten and unilever (all lowercase).

```{r Clean up brand names, echo=TRUE, message=FALSE, warning=FALSE}
# 1: Clean up brand names
# closeMatch function will replace the observations in the vector 
# with the closest match in the control vector
closeMatch <- function(x, y) {
  wt <- c()
    for (i in 1:length(y)) {
    wt[i] <- stringdist(x, y[i])
    }
  which.min(wt)
  return(y[which.min(wt)])
}
brandName <- c( "philips", "akzo", "van houten",  "unilever")
df$company <- sapply(df$company, closeMatch, y = brandName)
df
```

Separate the product code and product number into separate columns i.e. add two new columns called product_code and product_number, containing the product code and number respectively

```{r Separate product code and number, echo=TRUE, message=FALSE, warning=FALSE}
# 2: Separate product code and number
colnames(df)[which(names(df) == "Product.code...number")] <- "Productcode_number"
df <- separate(df, Productcode_number, c("Productcode", "number"), sep = "-", remove = FALSE)
df
```
You learn that the product codes actually represent the following product categories:
p = Smartphone
v = TV
x = Laptop
q = Tablet
In order to make the data more readable, add a column with the product category for each record.

```{r Add product categories, echo=TRUE, message=FALSE, warning=FALSE}
# 3: Add product categories
Productcode <- c("p", "v", "x", "q")
Productcategory <- c("Smartphone", "TV", "Laptop", "Tablet")
prodCatLookup <- data.frame(Productcode, Productcategory)
df <- left_join(df, prodCatLookup)
df
```

You'd like to view the customer information on a map. In order to do that, the addresses need to be in a form that can be easily geocoded. Create a new column full_address that concatenates the three address fields (address, city, country), separated by commas.

```{r Add full address for geocoding, echo=TRUE, message=FALSE, warning=FALSE}
# 4: Add full address for geocoding
full_address <- paste(df$address, df$city, df$country, sep = ", ")
df <- cbind(df, full_address)
df
```

Both the company name and product category are categorical variables i.e. they take only a fixed set of values. In order to use them in further analysis you need to create dummy variables. Create dummy binary variables for each of them with the prefix company_ and product_ i.e.,
* Add four binary (1 or 0) columns for company: company_philips, company_akzo, company_van_houten and company_unilever.
* Add four binary (1 or 0) columns for product category: product_smartphone, product_tv, product_laptop and product_tablet.

```{r Create dummy variables, echo=TRUE, message=FALSE, warning=FALSE}
# 5: Create dummy variables for company and product category
company <- gsub(" ", "_",df$company)
df2 <- dummy(company, sep = "_")
df <- cbind(df, df2)

product <- tolower(df$Productcategory)
df3 <- dummy(product, sep = "_")
df <- cbind(df, df3)
# View(df)
```

Save the output as a CSV file refine_original.csv

```{r Save output, echo=TRUE, message=FALSE, warning=FALSE}
write.csv(df, file = "refine_clean.csv")
```